---
title: "dataRefomatting"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r warning=FALSE, message=FALSE}
library(dplyr)
library(tidyr)

# these libraries have their own filter,lag functions. to call stats file, use   stats::filter()

```

# Reformat the Catch Data

* Remove the 'all' column
* create a species column
    -move from wide to long
* general QA

# Data

[data from: Mike Byerly. Alaska commercial salmon catches by management region (1886- 1997). urn:node:GOA. df35b.304.2.](https://knb.ecoinformatics.org/view/df35b.304.2)

```{r}

catch_original <- read.csv(url("https://knb.ecoinformatics.org/knb/d1/mn/v2/object/df35b.302.1", method = "libcurl"),
                    stringsAsFactors = FALSE)

head(catch_original)

```

# We will learn the following functions

* gather
* spread 
* rename
* mutate
* select
* filter
* group_by
* summarise
* arrange


Remove the 'all' and 'notesRegCode' columns using 'select'

hint:
*ctrl + shift + M : pipe shortcut %>%


```{r}
catch_long <- catch_original %>% 
    select(Region,Year,Chinook,Sockeye,Coho,Pink,Chum)



catch_long <- catch_long %>%  #select columns we want 
    gather(key="species", value="catch", Chinook, Sockeye, Coho, Pink, Chum)  # move from wde to long

head(catch_long)

```


or



{r eval=False}
catch_long <- catch_original %>% 
    select(-All, -notesRegCode)



catch_long <- catch_long %>% 
    gather(key="species", value="catch", -Region, -Year)


head(catch_long)

You can use the -Variable to exclude some rather than listing them all



```{r}
catch_wide <- catch_long %>% 
       spread(key=species, value =catch)

head(catch_wide)
```


# clean up data

* reaname 'Catch' to 'catch_thousands'
* create new catch column in units:  num. of fish
* hange catch column to be numerc


## rename

```{r}
catch_clean <-catch_long %>% 
    rename(catch_thousands = catch) %>%  # we are renaming the catch column
    mutate(catch_thousands=ifelse(catch_thousands=="I",1,catch_thousands)) %>% 
    mutate(catch_thousands=as.numeric(catch_thousands)) %>% 
    mutate(catch=catch_thousands*1000)  %>% 
    select(-catch_thousands)


head(catch_clean)

```

testing chunk - look for where as numeric failed, then go above and fix

{r}
 test <- as.numeric(catch_long$catch)
 i <- which (is.na(test)==TRUE)
 catch_long[i,]

## Split -  APPlY - cOMBINE

```{r}
species_mean <- catch_clean %>% 
     group_by(species) %>% 
     summarise(catch_mean=mean(catch),num_obs=n())
head(species_mean)
```


```{r}
species_mean <- catch_clean %>% 
     group_by(species, Region) %>% 
     summarise(catch_mean=mean(catch),num_obs=n())
head(species_mean)



```
```{r}
species_mean <- catch_clean %>% 
     group_by(species, Year) %>% 
     summarise(catch_mean=mean(catch),num_obs=n())
head(species_mean)
```

```{r}
species_mean <- catch_clean %>% 
     group_by(Year) %>% 
     summarise(catch_mean=mean(catch),num_obs=n()) %>% 
     arrange(-catch_mean)
head(species_mean)
```


#test filtering

```{r}
species_mean <- catch_clean %>% 
     group_by(Year) %>% 
     summarise(catch_mean=mean(catch),num_obs=n()) %>% 
     arrange(-catch_mean) %>% 
     filter(Year<= 1990 & Year >= 1960)

head(species_mean)
```



# Join the Region Definitions

read in regions dataframe

```{r}
region_defs <- read.csv(url("https://knb.ecoinformatics.org/knb/d1/mn/v2/object/df35b.303.1",
                            method = "libcurl"),
                        stringsAsFactors = FALSE) %>% 
    select(code, mgmtArea)

head(region_defs)

```



```{r}
catch_joined <- left_join(catch_clean, region_defs, by=c("Region"="code"))

head(catch_joined)
```

```{r}
catch_joined_INNER <- inner_join(catch_clean, region_defs, by=c("Region"="code"))

head(catch_joined_INNER)

```


# Misc functions: Seperate and Unite

## split

```{r}
dates_df <- data.frame(date = c("5/24/1930",
                                "5/25/1930",
                                "5/26/1930",
                                "5/27/1930",
                                "5/28/1930"),
                       stringsAsFactors = FALSE)


head(dates_df)
```

```{r}
dates_split <- dates_df %>% 
  separate(date, c("month", "day", "year"), sep="/", remove=F)

head(dates_split)
```
## Unite

```{r}
dates_unite <- dates_split %>% 
    unite(col="date_iso", year, month, day, sep="-")

head(dates_unite)

```

