---
title: "dataRefomatting"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r warning=FALSE, message=FALSE}
library(dplyr)
library(tidyr)

# these libraries have their own filter,lag functions. to call stats file, use   stats::filter()

```

# Reformat the Catch Data

* Remove the 'all' column
* create a species column
    -move from wide to long
* general QA

# Data

[data from: Mike Byerly. Alaska commercial salmon catches by management region (1886- 1997). urn:node:GOA. df35b.304.2.](https://knb.ecoinformatics.org/view/df35b.304.2)

```{r}

catch_original <- read.csv(url("https://knb.ecoinformatics.org/knb/d1/mn/v2/object/df35b.302.1", method = "libcurl"),
                    stringsAsFactors = FALSE)

head(catch_original)

```


Remove the 'all' and 'notesRegCode' columns using 'select'

hint:
*ctrl + shift + M : pipe shortcut %>%


```{r}
catch_long <- catch_original %>% 
    select(Region,Year,Chinook,Sockeye,Coho,Pink,Chum)



catch_long <- catch_long %>%  #select columns we want 
    gather(key="species", value="catch", Chinook, Sockeye, Coho, Pink, Chum)  # move from wde to long

head(catch_long)

```


or



{r eval=False}
catch_long <- catch_original %>% 
    select(-All, -notesRegCode)



catch_long <- catch_long %>% 
    gather(key="species", value="catch", -Region, -Year)


head(catch_long)

You can use the -Variable to exclude some rather than listing them all



```{r}
catch_wide <- catch_long %>% 
       spread(key=species, value =catch)

head(catch_wide)
```



# Join the Region Definitions





# Misc functions