---
title: "functiontest"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


# Work with some data in degrees

* CONVERT FROM f TO CELSIUS 
* wRITE A FUNCTION
* WRITE A FUNCTIONS  FROM c TO f
* dOCUMENT FUNCTION

```{r}
source('convert.R')

airtemps <- c(212, 100, 32, 64, 0, -20)

celsius1 <- (airtemps[1]-32)*5/9
celsius2 <- (airtemps[2]-32)*5/9
celsius3 <- (airtemps[3]-32)*5/9
celsius4 <- (airtemps[4]-32)*5/9
celsius5 <- (airtemps[5]-32)*5/9
celsius6 <- (airtemps[6]-32)*5/9

# or

celsius_all <- (airtemps-32)*5/9

new_fahr_to_celsius(airtemps)
```

# no wwrite a function to do it

```{r}

fahr_to_celsius <- function(fahr) {
  celsius <- (fahr-32)*5/9
  return(celsius)
}

celsius_temps=fahr_to_celsius(airtemps)
celsius_temps
```

## can define defualt values if the input is missing
```{r}

fahr_to_celsius <- function(fahr=212) {
  celsius <- (fahr-32)*5/9
  return(celsius)
}


fahr_to_celsius()

```



#  New function to go in revserse
```{r}
celsius_to_fahr <- function(cels) {
  fahr <- (cels*9/5)+32
  return(fahr)
}

celsius_to_fahr(100)


```
```{r}
celsius_to_fahr(celsius_temps)

```

# Documenting functions

```{r}
#' converts temperate in Degrees Fahrenheit to Celsius
#'
#' @param fahr for the input value (degrees)
#' @return the converted value in Celsius

fahr_to_celsius <- function(fahr) {
  celsius <- (fahr-32)*5/9
  return(celsius)
  
}
